---
title: "Final_Report"
author: "Cristian Sastre"
date: "12/09/2021"
output: 
        html_document:
         keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

The final report corresponding to the practical automatic learning course is shown below. In this project, data collected from accelerometers placed on 4 body parts of 6 participants was used in order to predict the way in which they performed the exercise. For the prediction, a model was trained (**Random forest**), which is then used to predict 20 cases obtained from the test set. 


# Libraries and Data

Loading required libraries and proposed data:
```{r results='hide', message=FALSE}
library(ggplot2)
library(kernlab)
library(rattle)
library(corrplot)
library(caret)

set.seed(1234)
```

```{r}
train_data <- read.csv("./pml-training.csv")
test_data <- read.csv("./pml-testing.csv")
dim(train_data)
dim(test_data)
```


# Cleaning the Data

Eliminating N / A variables and irrelevant variables:
```{r}
train_data <- train_data[,colMeans(is.na(train_data)) < .9]
train_data <- train_data[,-c(1:7)]
```

Removing near zero variance variables:
```{r nzv}
nvz <- nearZeroVar(train_data)
train_data <- train_data[,-nvz]
```

Division of the validation and training set:
```{r}
inTrain <- createDataPartition(y=train_data$classe, p=0.7, list=F)
train <- train_data[inTrain,]
valid <- train_data[-inTrain,]
```


# Correlation overview

Correlation matrix of variables in training set:
```{r}
corrPlot <- cor(train[, -length(names(train))])
corrplot(corrPlot, method="color")
```


# Train and Testing the Model

Training process for the random forest and results obtained in the confusion matrix:
```{r, cache=TRUE}
cont <- trainControl(method="cv", number=3, verboseIter=F) # 3-fold cross validation.
random_forest_model <- train(classe~., data=train, method="rf", trControl = cont, tuneLength = 5)
random_forest_prediction <- predict(random_forest_model, valid)
conf_matrix <- confusionMatrix(random_forest_prediction, factor(valid$classe))
conf_matrix
```

## Accuracy & Error

Summary of results, precision and error of the model: 
```{r, echo=FALSE}
accuracy <- round(c(conf_matrix$overall[1]),3) #accuracy
error <- 1 - accuracy #out of sample error
data.frame(accuracy = accuracy, error = error, row.names = "RF")
```


# Predictions on Test Set

Model prediction test, using the 20 cases given in the test set:
```{r}
print(predict(random_forest_model, test_data))
```
